---
title: "Normal Distribution and Interval Estimations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mlbench)
library(ggplot2)
```

```{r}
data(PimaIndiansDiabetes2)
df <- PimaIndiansDiabetes2
```

## Normal Distribution
### Generate Sample

```{r }
set.seed(1)

ns <- rnorm(n = 100, mean = 175, sd = 15)
ns
```

### Visual Check
```{r}
plot(ns)
```
```{r}
hist(ns, breaks = 20)
plot(density(ns, bw = 5))
```
```{r}
hist(ns, breaks = 20, freq = F)
lines(density(ns, bw = 5), col = "blue")
```
```{r}
mean(ns)
sd(ns)
n <- length(ns)
sd(ns) * sqrt(n-1) / sqrt(n)

library(MASS)
fitdistr(ns, "normal")
```

```{r}
hist(ns, breaks = 20, freq = F)
lines(density(ns, bw = 5), col = "blue")
lines(sort(ns), dnorm(sort(ns), mean = 176.6, sd = 13.4), col = "red")
```
### Formal Check

```{r}
shapiro.test(ns)
```

### Probabilities
```{r}
pnorm(q = 160, mean = 175, sd = 15)
pnorm(q = 160, mean = 175, sd = 15, lower.tail = F)
pnorm(q = 195, mean = 175, sd = 15, lower.tail = F)
pnorm(q = 195, mean = 175, sd = 15)

pnorm(q = 160, mean = 175, sd = 15) + pnorm(q = 195, mean = 175, sd = 15, lower.tail = F)

pnorm(q = 195, mean = 175, sd = 15) - pnorm(q = 160, mean = 175, sd = 15)

```

### TASK #2.1 Sigma Rules
- Check 1-2-3 sigma rules for normal distribution

## Central Limit Theorem

```{r}
set.seed(8)
population <- rnorm(1000000, 0, 1)
sample_10 <- sample(population, 10)
sample_50 <- sample(population, 50)
sample_100 <- sample(population, 100)

mean(sample_10)
mean(sample_50)
mean(sample_100)
```


```{r}
sample_means <- numeric(30)
set.seed(4)
for (i in 1:30) {
    sample_means[i] <- mean(sample(population, 10))
}
hist(sample_means)
```

```{r}
sample_means <- numeric(30)
set.seed(4)
for (i in 1:30) {
    sample_means[i] <- mean(sample(population, 100))
}
hist(sample_means)
```

### TASK #2.2 CLT
- Generate population (n = 1000000)
- Fix the size k for sample (k=10, k=100, k=1000)
- Sample k observations from the population for 50 times
- Calculate mean for each sample
- Then calculate mean and sd - of sample means
- Visualize for different k (for example, using faceting with common x-axis)
- Compare

## Confidence Intervals

```{r}
SE <- function(x) {sd(x) / sqrt(length(x))}

SE(sample_10)
SE(sample_50)
SE(sample_100)
```

```{r}
qt( (1-0.95)/2, df = 9)
mean(sample_10) - 2.262157*SE(sample_10)
t.test(sample_10)
```
```{r}
t.test(sample_100)$conf.int
```
```{r}
t.test(sample_100, conf.level = 0.99)$conf.int
```

### TASK #2.3 Statistical Report, part 2
- Create a function for CI output, format: [xx.xx%, xx.xx%]
- Add 95%-CI for your descriptive statistics as shown below
- Is's important to have CI after Mean itself
```{r}
source('utils.R')
numerical_stats_ci(df[, -c(1,9)])
```
