---
title: "Categoricals"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mlbench)
library(dplyr)
library(ggplot2)
```

```{r}
data(PimaIndiansDiabetes2)
df <- PimaIndiansDiabetes2
```

## Categoricals

### Binomial
```{r}
summary(df$mass)
# BMI: let's assume normal as <25
# cut()
df$bmi <- as.factor(ifelse(df$mass < 25, "Normal", "Abnormal"))
summary(df$bmi)
```

```{r}
vec_cat <- df$bmi
summary(vec_cat)
table(vec_cat)
```

```{r}
length(vec_cat) # aware of NA!
sum(table(vec_cat))
```



```{r}
x <- 106
n <- 757
x / n * 100
```

```{r}
# Exact Binomial Test
binom.test(x, n, conf.level = 0.9)
binom.test(x, n)
```

#### TASK #3.1 Categorical Output 1
- Create functions for categorical output for binomial factor
```{r}
source('/Users/valeria.bogdanova/Documents/private/itmo/utils.R')
freq_output(x, n)
ci_cat_output(x, n)

freq_output(n-x, n)
ci_cat_output(n-x, n)
```

### Multinomial
```{r}
# BMI: let's assume normal as <25, non-critical abnormal < 35, else - critical
df$bmi_multi <- as.factor(ifelse(df$mass < 25, "Normal",
                          ifelse(df$mass < 35, "NoSig Abnorm", "Sig Abnorm")))
summary(df$bmi_multi)
```

```{r}
vec_cat_m <- df$bmi_multi

tt <- table(vec_cat_m)
tt_p <- tt / sum(tt)
round(tt_p * 100, 2)
```
```{r}
library(MultinomialCI)

# Sison-Graz Method
tt_p
multinomialCI(tt, alpha = 0.05)

```


#### TASK #3.2 Categorical Output 2
- Create function for categorical output for multinomial factor
```{r}
ci_multicat_output(ft = tt, k = 1)
ci_multicat_output(ft = tt, k = 2)
ci_multicat_output(ft = tt, k = 3)
```

### Contingency Table = Cross table

```{r}
table(df$bmi, df$diabetes)
table(df$bmi_multi, df$diabetes)
```

```{r}
# Pearson's Chi-squared Test 
chisq.test(df$bmi, df$diabete)
chisq.test(table(df$bmi, df$diabetes))

chisq.test(df$bmi_multi, df$diabetes)
```
```{r}
# Fisher's Exact Test
# - for 2x2 cross tabs (but not only)
# - when for some cell count < 10 or <5 (you can meet different in literature)

fisher.test(df$bmi, df$diabete)

fisher.test(df$bmi_multi, df$diabete)
fisher.test(table(df$bmi_multi, df$diabetes))
```

#### TASK #3.3 Categorical Output Fully
- Study the example
- Perform categorical stats output both(!) for bmi and bmi_multi within the same report
```{r}
measure_order_cat <- c("N", "Missing",
                       "NormN", "NormF", "NormCi",
                       "AbnormN", "AbnormF", "AbnormCi")
df %>% select(bmi, diabetes) %>% 
    gather(key = 'Parameter', value = 'Value', -diabetes) %>% 
    group_by(diabetes, Parameter) %>% 
    summarise(N = n() - sum(is.na(Value)),
              Missing = sum(is.na(Value)),
              
              NormN = sum(Value == 'Normal', na.rm = T),
              NormF = freq_output(NormN, N),
              NormCi = ci_cat_output(NormN, N),
              
              AbnormN = sum(Value == 'Abnormal', na.rm = T),
              AbnormF = freq_output(AbnormN, N),
              AbnormCi = ci_cat_output(AbnormN, N)
    ) %>% ungroup() %>% 
    gather(key = 'Measure', value = 'Value', -c(diabetes, Parameter)) %>% 
    spread(key = diabetes, value = Value) %>% 
    arrange(match(Measure, measure_order_cat)) %>% 
    knitr::kable()

```






