---
title: "HW3_mrkd"
author: "aliya yakupova"
date: "12/15/2020"
output: html_document
---

```{r setup, include=FALSE}
#install.packages('stringr')
#install.packages('tidyverse')
#install.packages('maps')
library(tidyr)
library(dplyr)
library(stringr)
library(maps)
```

## R Markdown

We take database weather.rds

```{r weather, echo=FALSE}
weather <- readRDS("~/R_HW/weather.rds")
head(weather)
```

# Column-row transformation

After that with "gather" function we get the result, where columns become rows and we add key column called "day"

```{r weather_clean1, echo=FALSE}
weather_clean <- 
  weather %>% 
  gather(key = day, value = val, X1:X31, na.rm = TRUE)
weather_clean <- weather_clean[, -1]
weather_clean$day <- str_replace(weather_clean$day, "X", "")
head(weather_clean)
```

We spread measures to columns

```{r weather_clean3, echo=FALSE}
weather_clean  <- 
  weather_clean %>%
  spread(key = measure, value = val)
head(weather_clean)
```

# Date transformation

We unite dates with separator "-"

```{r weather_clean_date, echo=FALSE}
weather_clean  <- 
  weather_clean %>%
  unite(col = Date, year, month, day, sep = "-")
```

We also should convert column date to type date instead of character

```{r weather_clean7, echo=FALSE}
weather_clean  <- 
  weather_clean %>% 
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"))
str(weather_clean)
```

Now column Date has type "date"

# Type transformation

We see that for some columns with actual numerical data type is equal to character.

We should convert them to numeric.

```{r weather_clean6, echo=FALSE}
weather_clean  <- 
  weather_clean %>% 
  mutate_at(vars(CloudCover, Max.Dew.PointF:WindDirDegrees), funs(as.numeric))
```

We have error at PrecipitationIn because of value 'T', let's replace it to 10^(-5) and try to convert it again

```{r weather_clean_PrecipitationIn, echo=FALSE}
weather_clean$PrecipitationIn <- str_replace(weather_clean$PrecipitationIn, "T", "0.00001")#mutate

weather_clean  <- 
  weather_clean %>% 
  mutate(PrecipitationIn=as.numeric(PrecipitationIn))
```

Also let's convert Events to type "factor"

```{r weather_clean_events, echo=FALSE}
weather_clean  <- 
  weather_clean %>% 
  mutate_at(vars(Events), funs(as.factor))
```

As a result of type transformation we have
```{r weather_clean_type_res, echo=FALSE}
summary(weather_clean)
```

# Detecting NA

We see that columns Max.Gust.SpeedMPH and PrecipitationIn have 6 and 49 NA respectively.

```{r weather_clean_NA, echo=FALSE}
weather_clean[ which( is.na(weather_clean$Max.Gust.SpeedMPH) | is.na(weather_clean$PrecipitationIn)) , ] %>%
  select(Max.Gust.SpeedMPH, PrecipitationIn)
```

If we know what to do with such measurements we should replace them with data

# Detecting errors

We have Max.Humidity more then 100%

```{r weather_clean_humidity, echo=FALSE}
indexes <- which(weather_clean$Max.Humidity > 100)

weather_clean[indexes, ] %>% 
  select(Date, Events, Max.Humidity)
```

We decide that this is typo and change value to 100

```{r weather_clean_humidity_fix, echo=FALSE}
weather_clean$Max.Humidity[indexes] <- 100
```

Also we have negative Mean.VisibilityMiles
```{r weather_clean_miles, echo=FALSE}
indexes <- which(weather_clean$Mean.VisibilityMiles < 0)

weather_clean[indexes, ] %>% 
  select(Date, Events, Mean.VisibilityMiles)
```

Let's also assume that this is typo and change value to positive

```{r weather_clean_miles_fix, echo=FALSE}
weather_clean$Mean.VisibilityMiles[indexes] <- 1
```
  
# Finally

The final result of all transformations is below

```{r weather_clean_final, echo=FALSE}
summary(weather_clean)
```