```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mlbench)
library(ggplot2)
library(dplyr)
```


### TASK #2.1 Sigma Rules
- Check 1-2-3 sigma rules for normal distribution

```{r}
ns <- rnorm(n = 1000000)
mean_ <- mean(ns)
sd_ <- sd(ns)

calc_sigma <- function(ns_, mean_, sd_, num) {
  round((1 - (mean((ns_ > (mean_ + num * sd_))) + mean((ns_ < (mean_ - num * sd_))))) * 100, 2)
}

sprintf("Theoretical 1sigma = 68.26, checked value = %.2f", calc_sigma(ns, mean_, sd_, 1))
sprintf("Theoretical 2sigma = 95.44, checked value = %.2f", calc_sigma(ns, mean_, sd_, 2))
sprintf("Theoretical 3sigma = 99.72, checked value = %.2f", calc_sigma(ns, mean_, sd_, 3))
```

### TASK #2.2 CLT
- Generate population (n = 1000000)
- Fix the size k for sample (k=10, k=100, k=1000)
- Sample k observations from the population for 50 times
- Calculate mean for each sample
- Then calculate mean and sd - of sample means
- Visualize for different k (for example, using faceting with common x-axis)
- Compare

```{r}
population <- rnorm(1000000)
sample_size <- c(10, 100, 1000)

calc_means <- function(population_, k_, times_) {
  sample_means <- numeric(times_)
  for (t in 1:times_)
    sample_means[t] <- mean(sample(population_, k_))
  sample_means
}

par(mfrow=c(1,3))
for (k in sample_size) {
  sample_means <- calc_means(population, k, 50)
  mn <- mean(sample_means)
  std <- sd(sample_means)
  
  df <- data.frame(sample_means=sample_means, sample_size=k, mean=mn, std=std)
  print(ggplot(df, aes(x=sample_means)) + 
    geom_histogram(color = "orange", bins=30) + 
    geom_vline(xintercept = mn, color = "red", size = 0.4, show.legend = T) + 
    geom_vline(xintercept = std, color = "blue", size = 0.4, show.legend = T) + 
    geom_vline(xintercept = -std, color = "blue", size = 0.4, show.legend = T) +
    labs(title = sprintf("k=%.0f, mean=%.2f, sd=%.2f", k, mn, std)) +
    xlim(-1, 1))
}
```

### TASK #2.3 Statistical Report, part 2
- Create a function for CI output, format: [xx.xx%, xx.xx%]
- Add 95%-CI for your descriptive statistics as shown below
- Is's important to have CI after Mean itself

```{r}
SE <- function(x) {sd(x) / sqrt(length(x))}

CI <- function(x) {
    x <- x[!is.na(x)]
    vqt <- qt( (1-0.95)/2, df = 9)
    l <- round(mean(x) + vqt*SE(x), 2)
    u <- round(mean(x) - vqt*SE(x), 2)
    sprintf('[%.2f, %.2f]', l, u)
}
```

```{r}
data(PimaIndiansDiabetes2)
df <- PimaIndiansDiabetes2[2:8]

my_summary <- function(x) {
  q1 <- round(sapply(x, quantile, probs = 0.25, na.rm = TRUE), 2)
  q3 <- round(sapply(x, quantile, probs = 0.75, na.rm = TRUE), 2)
  res <- data.frame(
    N = sapply(x, function(x) sum(!is.na(x))),
    Missed = sapply(x, function(x) sum(is.na(x))),
    Mean = round(sapply(x, mean, na.rm = TRUE), 2), 
    '95% CI' = sapply(x, function(x) {
                                        x <- x[!is.na(x)]
                                        vqt <- qt( (1-0.95)/2, df = 9)
                                        l <- round(mean(x) + vqt*SE(x), 2)
                                        u <- round(mean(x) - vqt*SE(x), 2)
                                        sprintf("[%.2f, %.2f]", l, u)
                                    }),
    Median = round(sapply(x, median, na.rm = TRUE), 2),
    Min = round(sapply(x, min, na.rm = TRUE), 2),
    Max = round(sapply(x, max, na.rm = TRUE), 2),
    Q1 = q1,
    Q3 = q3,
    IQR = q3 - q1,
    SD = round(sapply(x, sd, na.rm = TRUE), 2)
    )
  
  res <- t(res)
  data.frame(res[,order(colnames(res))])
}

df <- my_summary(df)
tibble(cbind(Measurements = row.names(df), df))
```