```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mlbench)
library(dplyr)
library(ggplot2)
```


## TASK #4.1 Statistical Report with Tests Results
- Take any data set you like, containing 2 groups of observation (balance by size)
- Create report for numerical data with appropriate tests (description of 2 groups, comparing them)
- Create report for categorical data with appropriate tests (description of 2 groups, comparing them)
- When comparing you should put name of method and p-value

# Dataset choise and balance check
```{r}
dfi <- read.table('insurance.csv', header=T, sep=",")

nrow(dfi[dfi$sex == 'female',])
nrow(dfi[dfi$sex == 'male',])
```

# Numerical data
```{r}
# TODO auto-function
numericals <- names(dfi)[c(1,3,4,7)]
for (param in numericals) {
    p <- ggplot(dfi, aes(x = get(param), fill = sex)) +
        geom_density(alpha=0.5) +
        xlab(param) + theme_bw()
    print(p)
}
```

```{r}
# Normality test
shi <- shapiro.test(dfi$bmi[dfi$sex == 'female'])
paste(shi$method, ": p-value=", round(shi$p.value, 4), sep='')
shi <- shapiro.test(dfi$bmi[dfi$sex == 'male'])
paste(shi$method, ": p-value=", round(shi$p.value, 4), sep='')

#distributions are not normal, we can't apply t-test
```

```{r}
mwi <- wilcox.test(bmi ~ sex, data = dfi)
paste(mwi$method, ": p-value=", round(mwi$p.value, 4), sep='')
# We can't reject H0, samples can come from same population
```
# Automatization
```{r}
pval_check <- function(pvalue){
  if (pvalue > 0.05) { "can't reject" }
  else { "reject" }
}

numerical_test <- function(df, var1, var1_name, var2, var2_name, vals){
 
  Objects <- c('')
  Type <- c('')
  Normal_check <- c('')
  Method <- c('')
  P_value <- c('')
  H0 <- c('') 
  
  shi1 <- shapiro.test(var1[var2 == vals[1]])
  shi2 <- shapiro.test(var1[var2 == vals[2]])
  shi_test <- (round(shi1$p.value, 4) >= 0.05) & (round(shi2$p.value, 4) >= 0.05)
  
  fisher <- var.test(var1 ~ var2, data = df)
  fisher_test <- round(fisher$p.value, 4) >= 0.05
  
  Objects <- append(Objects, paste(var1_name, ' vs ', var2_name))
  Type <- append(Type, 'Numerical')
  
  if ((shi_test) & (fisher_test)) {
    tt <- t.test(var1 ~ var2, data=df, var.equal = T)
    pvalue <- round(tt$p.value, 4)
    
    Normal_check <- append(Normal_check, '+')
    Method <- append(Method, tt$method)
    P_value <- append(P_value, toString(pvalue))
    H0 <- append(H0, pval_check(pvalue))
  }
  else {
    mwi <- wilcox.test(var1 ~ var2, data = df)
    pvalue <- round(mwi$p.value, 4)
    
    Normal_check <- append(Normal_check, '-')
    Method <- append(Method, mwi$method)
    P_value <- append(P_value, toString(pvalue))
    H0 <- append(H0, pval_check(pvalue))
  }

  res <- data.frame(Objects, Type, Normal_check, Method, P_value, H0)
  return(res)
}

res <- numerical_test(dfi, dfi$bmi, 'bmi', dfi$sex, 'sex', c('female', 'male'))
res %>% knitr::kable()
```
```{r}
s1 <- c('1','2')
s1_ <- append(s1, '3')
s2 <- c('4','5')
s2_ <- append(s2, '6')
res <- data.frame(s1_, s2_)
res %>% knitr::kable()
```
# Categorical data
```{r}
# TODO auto-function
cti <- chisq.test(dfi$smoker, dfi$sex)
paste(cti$method, ": p-value=", round(cti$p.value, 4), sep='')

fti <- fisher.test(dfi$smoker, dfi$sex)
paste(fti$method, ": p-value=", round(fti$p.value, 4), sep='')

#fti$estimate
#fti$conf.int
```

# Automatization
```{r}
pval_check <- function(pvalue){
  if (pvalue > 0.05) { "can't reject" }
  else { "reject" }
}

categorical_test <- function(var1, var1_name, var2, var2_name){
 
  Objects <- c('')
  Type <- c('')
  Method <- c('')
  P_value <- c('')
  H0 <- c('')
  
  Objects <- append(Objects, paste(var1_name, ' vs ', var2_name))
  Type <- append(Type, 'Categorical')
  
  if (length(cbind(var1, var2)) > 4) {
    cti <- chisq.test(var1, var2)
    pvalue <- round(cti$p.value, 4)
    
    Method <- append(Method, cti$method)
    P_value <- append(P_value, toString(pvalue))
    H0 <- append(H0, pval_check(pvalue))
  }
  else {
    fti <- fisher.test(var1, var2)
    pvalue <- round(fti$p.value, 4)
    
    Method <- append(Method, fti$method)
    P_value <- append(P_value, toString(pvalue))
    H0 <- append(H0, pval_check(pvalue))
  }

  res <- data.frame(Objects, Type, Method, P_value, H0)
  return(res)
}

res <- categorical_test(dfi$smoker, 'smoker', dfi$sex, 'sex')
res %>% knitr::kable()
```
# Result
```{r}
Objects=c('bmi vs sex', 'smoker vs sex', 'smoker vs sex')
Type=c('Numerical', 'Categorical', 'Categorical')
Method=c(mwi$method, cti$method, fti$method)
P_value=c(round(mwi$p.value, 4), round(cti$p.value, 4), round(fti$p.value, 4))
H0=c("can't reject", "reject", "reject")

res <- data.frame(Objects, Type, Method, P_value, H0)
res %>% knitr::kable()
```
