---
title: "Comparing Two Groups"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mlbench)
library(dplyr)
library(ggplot2)
```

```{r}
data(PimaIndiansDiabetes2)
df <- PimaIndiansDiabetes2
```

## Comparing Numericals
### Visual Check
```{r}
numericals <- names(df)[2:8]
for (param in numericals) {
    p <- ggplot(df, aes(x = get(param), fill = diabetes)) +
        geom_density(alpha=0.5) +
        xlab(param) + theme_bw()
    print(p)
}

```

### Parametric: Student's t-Test
```{r}
# 1) check both samples for normal distribution
# 2) variances are homogeneous (if not - Welch modification)
# 3) perform 2-sample t-test

# Normality test
sh <- shapiro.test(df$pressure[df$diabetes == 'neg'])
sh
sh <- shapiro.test(df$pressure[df$diabetes == 'pos'])
sh

str(sh)

sh$method
round(sh$p.value, 4)
```

```{r}
# Fisher test for variances
var.test(pressure ~ diabetes, data = df)
```

```{r}
# actually, we are not eligible to do so in given situation! but...
# var.equal = FALSE
tt <- t.test(df$triceps[df$diabetes == 'pos'],
             df$triceps[df$diabetes == 'neg'])
tt

tt <- t.test(triceps ~ diabetes, data=df, var.equal = T)
tt

str(tt)
tt$method
round(tt$p.value, 4)
round(tt$estimate, 2)
tt$conf.int
```

### Non-Parametric:  Wilcoxon test / Mann-Whitney test

```{r}
# actually, we should use this approach =)
mw <- wilcox.test(df$triceps[df$diabetes == 'pos'],
                  df$triceps[df$diabetes == 'neg'])
mw

mw <- wilcox.test(triceps ~ diabetes, data = df)
mw

str(mw)
mw$method
round(mw$p.value, 4)
```

### Paired Tests

```{r}
# t.test(..., paired = FALSE -> paired = TRUE)
# wilcox.test(..., paired = FALSE -> paired = TRUE)
```

## Comparing Categoricals

```{r}
df$bmi <- as.factor(ifelse(df$mass < 25, "Normal", "Abnormal"))
ct <- chisq.test(df$bmi, df$diabete)
ct

str(ct)
ct$method
ct$p.value


ft <- fisher.test(df$bmi, df$diabete)
ft
str(ft)
ft$method
ft$p.value
ft$estimate
ft$conf.int
```

### Paired tests

```{r}
# mcnemar.test() # 2x2
```

## TASK #4.1 Statistical Report with Tests Results
- Take any data set you like, containing 2 groups of observation (balance by size)
- Create report for numerical data with appropriate tests (description of 2 groups, comparing them)
- Create report for categorical data with appropriate tests (description of 2 groups, comparing them)
- When comparing you should put name of method and p-value
### Example will be attached later
