---
title: "R 1st term project"
author: "aliya yakupova"
date: "1/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(RColorBrewer)
```

## Dataset descripton

For the project we will take database of **locations of Earthquakes off Fiji** (<https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/quakes.html>)

The data set give the locations of 1000 seismic events of MB > 4.0. The events occurred in a cube near Fiji since 1964.

## Format

A data frame with 1000 observations on 5 variables.

1.	lat	numeric	**Latitude of event**

2.	long	numeric	**Longitude**

3.	depth	numeric	**Depth (km)**

4.	mag	numeric	**Richter Magnitude**

5.	stations	numeric	**Number of stations reporting**

```{r pressure, echo=FALSE}
# plot(pressure)
require(graphics)
quakes %>%
  select(3:5) %>%
  pairs(main = "Fiji Earthquakes, N = 1000", cex.main = 1.2, pch = ".")
```

```{r quakes}
str(quakes)
summary(quakes)
sum(is.na(quakes))
```

As we see there are no NA and all types are proper.

```{r quakes_plot, echo=FALSE}
quakes %>%
  #mutate(magnitude = rev(mag)) %>%
  ggplot(aes(x = long,
             y = lat,
             color = mag)) +
  geom_point() +
  scale_fill_distiller(palette ="RdBu", direction = -1)
  #scale_color_manual(values = rev(RColorBrewer::brewer.pal(3,'Blues'))) 
  #geom_point(data = filter(quakes, mag >= 6), color = "red")
  #scale_fill_discrete(name = "Remarkable", labels = c("mag >= 6"))
```

We can see that mostly earthquakes took a place at longitude 176 -- 186 and latitude -30 -- -15

Now let's have a look where were the remarkable strength of earthquakes by Richter scale. 

Magnitude 5 is known as moderate and higher is scaled like strong. At the plot earthquakes with magnitude equal or higher than 6 are colored red.

As we can see there are very few strong earthquakes with remarkable damage.

Now let's have a look at correlation of magnitude and depth.

```{r quakes_plot_mag_depth, echo=FALSE}
quakes %>%
  ggplot(aes(x = mag,
             y = depth,
             color = stations)) +
  geom_point()
```

We can't say that there is no significant correlation.

Now have a look how many stations detected earthquake respectively to depth, separately for remarkable earthquakes and not according to magnitude bigger or equal to 5.

```{r quakes_stations, echo=FALSE}
quakes %>% 
  select(depth,mag,stations) %>%
  mutate(strong = ifelse(mag >= 5, "remarkable magnitude", "not remarkable")) %>%
  ggplot(aes(x = stations,
             y = depth,
             color = stations)) +
  geom_point() + 
  geom_point(size = 0.7) +
  facet_grid(strong ~ ., scales="free")
```

To see how depth and magnitude correlates with detection of earthquake by station let's plot their dependency.

```{r quakes_gather, echo=FALSE}
quakes %>% 
  gather(key = measure, value = value, "depth":"mag") %>%
  ggplot(aes(x = stations,
             y = value,
             color = stations)) +
  geom_point() + 
  geom_point(size = 0.7) +
  facet_grid(measure ~ ., scales="free")
```

And we see that the higher magnitude the higher chance for detecting and with depth it doesn't work.