---
title: "Descriptive Statistics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mlbench)
library(ggplot2)
```

## Basics
### Data Overview
```{r}
data(PimaIndiansDiabetes2)
df <- PimaIndiansDiabetes2
#View(df)

str(df)
```



```{r}
summary(df)
```

### Descriptive Statistics


```{r}
vec <- df$mass
summary(vec)
```
```{r}
min(vec); max(vec)
mean(vec); median(vec)
```
```{r}
min(vec, na.rm = T); max(vec, na.rm = T)
mean(vec, na.rm = T); median(vec, na.rm = T)
```


```{r}
quantile(vec, na.rm = T)
# quantile(vec, probs = 0.5)
quantile(vec, probs = c(0.25, 0.75), na.rm = T)
```
```{r}
length(vec)

sum(is.na(vec))
```
```{r}
sd(vec, na.rm = T)
var(vec, na.rm = T)

sqrt(var(vec, na.rm = T))
sd(vec, na.rm = T)^2
```
```{r}
vec_clean <- na.omit(vec)
sd(vec_clean)
```
```{r}
range(vec_clean)
IQR(vec_clean)
```

#### TASK #1.1 Measures of Center
- Write your own function or one-liner to calculate mean, median, mode
- Check the correctness using built-in functions for mean and median (there is no function for mode - try your one for different cases)
- Visualize (histogram or density plot) the result for all these measures
- Spoil you data with an outlier and repeat previous step
#### Example
```{r}
x <- c(175, 176, 182, 165, 167, 172, 175, 196, 158, 172)
x_mean <- mean(x)
x_sp <- c(x, 210)
x_mean_sp <- mean(x_sp)

original <- data.frame(height = x, group = "original")
spoiled <- data.frame(height = x_sp, group = "spoiled")
dataviz <- rbind(original, spoiled)

vline <- data.frame(z = c(x_mean, x_mean_sp), group = c("original", "spoiled"))

ggplot(dataviz, aes(x = height)) +
    geom_histogram(binwidth = 5) +
    geom_vline(aes(xintercept = z), vline, colour = "red")+
    facet_grid(group ~ .) +
    theme_bw() + ggtitle("Mean")
```

#### TASK #1.2 Measures of Spread
- Write your own function or one-liner to calculate variance and sd
- Check the correctness using built-in functions 
- Visualize (boxplot) the result for: Q1+Q3, 1-sd borders
- Spoil you data with an outlier and repeat previous step
#### Example
```{r}
x_range <- range(x)
x_range_sp <- range(x_sp)

hline <- data.frame(z = c(x_range, x_range_sp),
                    group = rep(c("original", "spoiled"), each = 2))

ggplot(dataviz, aes(y = height)) +
    geom_boxplot() +
    geom_hline(aes(yintercept = z), hline, colour = "red")+
    facet_grid(. ~ group) +
    theme_bw() + ggtitle("Range")
```
### Properties
```{r}
mean(x - 100); mean(x) - 100 # 1
mean(x / 100); mean(x) / 100 # 2
sum(x - mean(x)) # 3
```
#### Example
```{r}
x_ms_100 <- x - 100

original <- data.frame(height = x, group = "original")
trans_ms <- data.frame(height = x_ms_100, group = "x - 100")
dataviz <- rbind(original, trans_ms)

x_ms_mean <- mean(x_ms_100)
vline <- data.frame(z = c(x_mean, x_ms_mean), group = c("original", "x - 100"))

ggplot(dataviz, aes(x = height)) +
    geom_density() +
    geom_vline(aes(xintercept = z), vline, colour = "red")+
    facet_wrap(vars(group), nrow = 2, scales = 'free_y') +
    theme_bw() + ggtitle("Mean Properties")
```
```{r}
var(x - 100); var(x) # 4
var(x / 100); var(x) / 10000 # 5
sd(x / 100); sd(x) / 100 # 6
```

#### TASK #1.3 Properties of Measures
- Check the properties for your sample
- Visualize result tabularly (for # 1-6) and graphically (for # 2 and 6)

### Custom Report
#### TASK #1.4 Statistical Report, part 1
Create descriptive statistics on Pima data set (excluding first and last column)
containing information for each column such as:
- Number of valid observations (not NA)
- Number of missing (NA) values
- Mean, Median
- Min, max
- Q1, Q3
- IQR, SD

Order of columns:
- Measure column first, then all the others in free order
Order of measures:
- As presented below

Values rounded to the 2nd digit
```{r}
source('utils.R')
numerical_stats(df[, -c(1,9)])
```


