```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mlbench)
library(dplyr)
library(tidyr)
library(ggplot2)
library(corrplot)
```


## TASK 5

Anscombe’s data set
• Scatter plot facetted by set
• Summary calculation (mean, sd) grouped by set
• Pearson’s correlation by set, and non-parametric, and p-values
• Add geom_smooth() to the plot

```{r}
data(anscombe)
raw_df <- anscombe
raw_df %>% knitr::kable()
```

```{r}
X = gather(raw_df[1:4])
Y = gather(raw_df[5:8])
df <- data.frame(X=X$value, 
                 Y=Y$value, 
                 XY=paste0('(', X$key, ',', Y$key, ')'))
head(df) %>% knitr::kable()
```

```{r}
p <- ggplot(df, aes(X, Y)) + 
  geom_point() +
  facet_wrap(~XY, ncol = 2) +
  geom_smooth(formula = y ~ x, method = "lm", fullrange = T)
p
```

```{r}
colnames <- c("mean X", "mean Y", 'sd X','sd Y','Pearson correlation','Kendall correlation', 'Spearman correlation','p-value')

summary <- df %>% 
    group_by(XY) %>% 
    summarize("mean X" = round(mean(X, na.rm = T), 4),
              "mean Y" = round(mean(Y, na.rm = T), 4),
              "sd X" = round(sd(X, na.rm = T), 4),
              "sd Y" = round(sd(X, na.rm = T), 4),
              "Pearson correlation" = round(cor(X, Y, method = "pearson"), 4),
              "Kendall correlation" = round(cor(X, Y, method = "kendall"), 4),
              "Spearman correlation" = round(cor(X, Y, method = "spearman"), 4),
              "p-value" = signif(cor.test(X, Y)$p.value, 1)) %>%  
    gather(key = 'Feature', value = 'Value', -c(XY))  %>%  
    spread(key = XY, value = Value) %>% 
    arrange(match(Feature, colnames))

summary %>% knitr::kable()
```

# TASK 6

Other data set: https://archive.ics.uci.edu/ml/datasets/Air+quality
• Explore data set, clean if needed
• Explore each variable independently
• Cross correlations
• Build simple linear models with each predictor, check assumptions
• For one of the models create train-test sets, plot the model, for the test set color
real and predicted points differently; R^2 and p-value to title

```{r}
df <- read.table("AirQualityUCI/AirQualityUCI.csv", sep = ";", header = T, dec = ",")
summary(df)
```
```{r}
dim(df)
```

```{r}
df <- na.omit(df[, -c(1, 2, 16, 17)])
df[df == -200] <- NA
#for(i in 1:ncol(df)){
#  df[is.na(df[,i]), i] <- mean(df[,i], na.rm = TRUE)
#}
summary(df)
```
```{r}
dim(df)
```
  
```{r}
head(df) %>% knitr::kable()
```

```{r}
my_summary <- function(df) {
      mean <- sapply(df, function(x) round(mean(x, na.rm = T), 2))
      median <- sapply(df, function(x) round(median(x, na.rm = T), 2))
      min <- sapply(df, function(x) round(min(x, na.rm = T), 2))
      max <- sapply(df, function(x) round(max(x, na.rm = T), 2))
      CI <- sapply(df, function(x) paste(round(t.test(x)$conf.int, 2), collapse = " : "))
      Std <- sapply(df, function(x) round(sd(x, na.rm = T), 2))
      Q1 <- sapply(df, function(x) round(quantile(x, na.rm = T)[c(2)], 2))
      Q3 <- sapply(df, function(x) round(quantile(x, na.rm = T)[c(4)], 2))
      IQR <- round((Q3 - Q1), 2)
    
      res <- data.frame(Mean=mean, Median=median, Min=min, Max = max,
                        Q1=Q1, Q3 = Q3, IQR = IQR)
      return(res)
}

my_summary(df) %>% knitr::kable()
```
```{r}
pairs(df)
```

```{r}
dfg <- gather(df)

p <- ggplot(dfg, aes(key, value)) + 
  geom_boxplot(notch=TRUE)

p + facet_wrap(~key, ncol = 5 , scales = "free") + 
  theme_light()
```
```{r}
corrplot.mixed(cor(df, use="pairwise.complete.obs", method = 'spearman'),
               lower = "number", 
               upper = "circle",
               tl.col = "black", 
               tl.pos = "lt",
               sig.level = 0.05, 
               insig = "blank")
```
```{r}
# Function to add correlation coefficients
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...) {
    usr <- par("usr")
    on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    Cor <- abs(cor(x, y, method = 'spearman', use="pairwise.complete.obs")) # Remove abs function if desired
    txt <- paste0(prefix, format(c(Cor, 0.123456789), digits = digits)[1])
    if(missing(cex.cor)) {
        cex.cor <- 0.4 / strwidth(txt)
    }
    text(0.5, 0.5, txt,
         cex = 1 + cex.cor * Cor) # Resize the text by level of correlation
}

# Plotting the correlation matrix
pairs(df,
      upper.panel = panel.cor,    # Correlation panel
      lower.panel = panel.smooth) # Smoothed regression lines
```
```{r}
#install.packages("GGally")
#library("GGally", verbose = FALSE, quietly = TRUE)
#d <- df[3:15]
#p_ <- GGally::print_if_interactive
#pm <- ggpairs(d, 
#        columns = colnames(d), 
#        upper = list(continuous = wrap("cor", size = 10)), 
#        lower = list(continuous = "smooth"))
#print(pm)
```
```{r}
#d <- df[3:15]
#d %>% ggpairs(., 
#               lower = list(continuous = wrap("smooth", alpha = 0.3, size=0.1)))
```

```{r}
storage <- list()
colpairs <- t(combn(names(df), 2))
for(i in 1:(length(colpairs)/2)){
  storage[[i]] <- lm(get(colpairs[i,1]) ~ get(colpairs[i,2]), data=df)
  plotname <- paste(colpairs[i,1],"~",colpairs[i,2])
  print(summary(storage[[i]]))
  print(plot(storage[[i]], which = c(1,2), sub=plotname))
}
```

```{r}
#lm <- lm(PT08.S2.NMHC. ~ PT08.S1.CO., data=df)
#summary(lm)
#plot(lm, which = c(1,2))
```

```{r}
set.seed(88)
data <- df
sample <- sample.int(n = nrow(data), size = floor(.75*nrow(data)))
train <- data[sample, ]
test <- data[-sample, ]
new_mod <- lm(data = train, CO.GT. ~ C6H6.GT.)
summary(new_mod)
```
```{r}
p <- train %>% ggplot(aes(x = C6H6.GT., y = CO.GT.)) + geom_point() + geom_smooth(method = "lm")
print(p)
```

```{r}
pred <- predict(new_mod, newdata = test)
head(pred)
```

```{r}
test$CO.GT._pred <- pred
head(test[c("CO.GT.","CO.GT._pred")])
```

```{r}
df_train <- data.frame(key='train', "CO.GT."=train$CO.GT., "C6H6.GT."=train$C6H6.GT.)
df_test <- data.frame(key='test', "CO.GT."=test$CO.GT., "C6H6.GT."=test$C6H6.GT.)
df_predicted <-data.frame(key='predicted', "CO.GT."=test$CO.GT._pred, "C6H6.GT."=test$C6H6.GT.)

long <- rbind(df_train, df_test, df_predicted)
              
head(long)
```

```{r}
p <- long %>%
  ggplot(aes(x = C6H6.GT., y = CO.GT., col=key)) + 
  geom_point() +
  geom_smooth(method = "lm", se=F)

R2 <- round(summary(new_mod)$r.squared*100,2)
pvalue <- summary(new_mod)$coefficients[2,4]

p + ggtitle(paste("R^2 :", R2, ", p-value : ", pvalue))

```
# MULTIPLE LINEAR REGRESSION

CO.GT. - True hourly averaged concentration CO in mg/m^3 (reference analyzer)
C6H6.GT. - True hourly averaged Benzene concentration in microg/m^3 (reference analyzer)
NO2.GT. - True hourly averaged NO2 concentration in microg/m^3 (reference analyzer)
PT08.S4.NO2. - PT08.S4 (tungsten oxide) hourly averaged sensor response (nominally NO2 targeted)

```{r}
new_mod10 <- lm(data = train, CO.GT. ~ C6H6.GT.)
summary(new_mod10)
```

```{r}
new_mod11 <- lm(data = train, CO.GT. ~ NO2.GT.)
summary(new_mod11)
```

```{r}
new_mod12 <- lm(data = train, CO.GT. ~ PT08.S4.NO2.)
summary(new_mod12)
```

```{r}
new_mod2 <- lm(data = train, CO.GT. ~ C6H6.GT. + NO2.GT.)
print(summary(new_mod2))
print(plot(new_mod2))
```
```{r}
pred2 <- predict(new_mod2, newdata = test)
test$CO.GT._pred2 <- pred2
head(test[c("CO.GT.","CO.GT._pred2")])
```

```{r}
print(
  paste0(
    "R^2: ", round(summary(new_mod2)$r.squared*100,2),
    ", pvalue: C6H6.GT.=", summary(new_mod2)$coefficients[2,4], "; NO2.GT.=", summary(new_mod2)$coefficients[3,4]
  )
)
```

```{r}
new_mod3 <- lm(data = train, CO.GT. ~ C6H6.GT. + NO2.GT. + PT08.S4.NO2.)
print(summary(new_mod3))
print(plot(new_mod3))
```

```{r}
pred3 <- predict(new_mod3, newdata = test)
test$CO.GT._pred3 <- pred3
head(test[c("CO.GT.","CO.GT._pred3")])
```
```{r}
print(
  paste0(
    "R^2: ", round(summary(new_mod3)$r.squared*100,2),
    ", pvalue: C6H6.GT.=", summary(new_mod3)$coefficients[2,4], 
            "; NO2.GT.=", summary(new_mod3)$coefficients[3,4], 
            "; PT08.S4.NO2=", summary(new_mod3)$coefficients[4,4]
  )
)
```